---
title: "Getting started with sitingclass"
author: "Pierre-Marie Lefeuvre"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{getting_started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r init,echo=FALSE}
knitr::opts_chunk$set(collapse=FALSE, #True
                      comment=NA, #"#>"
                      warning=FALSE,
                      message=FALSE,
                      cache=FALSE,
                      verbose=TRUE,
                      fig.width=4,
                      fig.height=4,
                      fig.align='center',
                      fig.cap='center',
                      out.width='50%',
                      dev='jpeg',
                      dev.args=list(quality=25,
                                    type='cairo'))
```

# Getting started with `sitingclass`
The `sitingclass` package computes the exposure of a weather station. Currently, the package is built for Norwegian weather stations that are maintained by the Norwegian Meteorological Institute (https://www.met.no/). 

Check the README for information on this R package's installation, authentication, dependencies, as well as the source of used data and metadata.

From R, get the following examples by loading the library and running: 
```{r, eval=FALSE}
vignette("getting_started", package = "sitingclass")
```

# Plot the station with background imagery
## Get the station's metadata
Load the library and extract the metadata for a given station using the Met's stationid

```{r}
library(sitingclass)

# Define station ID as used by other functions below
stationid <- 18700

# Get station metadata
stn <- get_metadata_frost(stationid, path = NULL)
# Set default path to NULL in order to print and not save plots
```

The result is a SpatVector with coordinates in UTM33 (unit: metre) and metadata: 

* the station: name, coordinates (lat/lon), alternative ids, exposure, organisation, 
* the physical parameter: name, id, level, sensor, performance,
* the processing parameters: 
  + `dx`: the radius in metre (default: 1) of the boundary box that defines the area to download
  + `resx`: the spatial resolution in metre (default: 100)for the data to download
  + `path`: where to save the output figures (default: output/_stationid_)

```{r}
stn
# Reverse dataframe to show last added parameters
rev(as.data.frame(stn))
```
## Construct a boundary box

### From SpatVector
The easiest method is to use the default processing parameters that are contained in the SpatVector. One could edit their values directly during loading of the metadata with:
`get_metadata_frost(stationid, dx=1, resx=100, path="output/stationid"`

```{r}
box <- make_bbox(stn)
box
```

### From coordinates

The other method is to manually extract the coordinates from the SpatVector and set manually the radius `dx`
```{r}
# Get coordinates
centre <- terra::crds(stn)
centre

# Define radius
dx <- 100

# Construct box to extract WMS tile
box <- make_bbox(centre, dx = dx)
box
```

## Plot background tile with station

The boundary box is then used to get background maps from WMS layers, in this case `"esri"` satellite imagery. The list of all available layers is found in the function documentation.
```{r, fig.cap="A satellite image with the station location as a red dot."}
# Plot ESRI imagery tile
plot_tile_station(stn, box, tile_name = "esri")
```

## Plot station, tile and grid
The function `plot_station_grid()` adds a grid and circles to the image. The grid/circle size and interval are defined by the radius of the image with default 10-m, 50-m, 100-m and 1000-m.

In this example, the tile name `"ortofoto"` gets aerial imagery from _Kartverket_, the Norwegian Mapping Authority. 
```{r, fig.cap="Grid and buffer with 'Norge i bilder' imagery."}
# Plot tile with grid and proximity circles
plot_station_grid(stn, tile_name = "ortophoto_demo", grid_scale = 10)
```

# Plot Digital Elevation Models (DEM)
## Download DEMs from Kartverket
The area of the DEM is defined by the boundary box set above in addition to the spatial resolution `resx` that is required by Kartverket's Web Coverage Service (WCS). 

Two types of models are available:

* DTM: Digital Terrain Model that represents the ground elevation (`"dtm"` in Norwegian)
* DSM: Digital Surface Model that represents the surface elevation including trees and buildings (`"dom"` in Norwegian)

The function fetches already downloaded and stored locally DTMs if the flag `f_overwrite` is set to `FALSE`, otherwise it overwrites the file, the default behaviour.
```{r, results = 'hide'}
f_ow  <- FALSE
dem   <- download_dem_kartverket(stn,
                                 name = "dtm",
                                 f_overwrite = f_ow)
dsm   <- download_dem_kartverket(stn,
                                 name = "dom",
                                 f_overwrite = f_ow)
```

## Plot DEMs with `rayshader`
The package `rayshader` is used here to produce 2D and 3D visualisations of the DEMs with shaded relief. Here is an example where the 100-m DSM raster is converted to a matrix and then plotted with a shaded relief.
```{r, fig.cap="Hillshade of the 100-m digital surface model."}
library(rayshader)
elmat   <- raster_to_matrix(dsm)
elmat %>%
  sphere_shade(texture = "desert") %>%
  plot_map()
```

## Example for larger areas
For larger areas, one can also set the boundary box manually in `download_dem_kartverket()` with `stn` being the centre, `dx` the radius of the boundary box and `resx` the horizontal resolution of the model. In this example, a  20-km radius, 20-m resolution DEM is downloaded and plotted.
```{r, fig.cap="Hillshade of the 20-km digital elevation model.", results = 'hide'}
demkm <- download_dem_kartverket(stn,
                                 name = "dtm",
                                 dx = 20e3,
                                 resx = 20,
                                 f_overwrite = f_ow)
elmatkm <- raster_to_matrix(demkm)
elmatkm %>%
  sphere_shade(texture = "desert") %>%
  plot_map()
```

# Plot a sun diagram
The function `plot_station_horizon_sun()` computes and plots sun diagram from the following inputs:

* the station location (`stn`)
* a local and regional Digital Elevation Model  (`dem` and `demkm`) to show shading from the landscape
* a local Digital Surface Model (`dsm`) to show shading from local buildings or trees. 

```{r, fig.dim=c(6,4), fig.cap=paste("Horizon and sun diagram for the station ",stationid), results = 'hide'}
plot_station_horizon_sun(stn)
```

# Compute the final siting class
The function `plot_station_siting_context()` eventually contains a workflow producing the figures necessary for station inspections and estimate the final siting class:

* Gets the metadata of a Met weather station
* Plot station with grid and circles covering the area with the radii 10-m, 50-m, 100-m, 1000-m
* Plot the sun diagram
* Compute the siting class of the station
```{r, eval=FALSE}
# Set timezone to avoid time shift between winter and summer time
Sys.setenv(TZ = "UTC") # "Europe/Oslo"

# The main function plotting sun diagram and context for a weather station
plot_station_siting_context(stationid = 18700,
                            paramid = 211,
                            f_verbose = FALSE,
                            f_pdf = FALSE)
```


