---
title: "sitingclass"
author: "Pierre-Marie Lefeuvre"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sitingclass}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# `sitingclass`
The `sitingclass` package computes the exposure of a weather station. Currently, the package is built for Norwegian weather stations that are maintained by the Norwegian Meteorological Institute (aka Met). 

## Startup
### Installation
The package `sitingclass` is available on Met's Gitlab and can be installed using `remotes::install_git()`. Send an access request to [pierre.marie.lefeuvre@met.no](mailto:pierre.marie.lefeuvre@met.no) to get `YOUR_GIT_ID` and `YOUR_GIT_TOKEN`. See [Authentication](#Authentication) on how to make your id and token accessible to your R environment.

```{r, eval=FALSE}
# manually adding your token
remotes::install_git("https://YOUR_GIT_ID:YOUR_GIT_TOKEN@gitlab.met.no/pierreml/sitingclass.git",force=T)

# If you have set your token in .Renviron, use:
remotes::install_git(paste0("https://",Sys.getenv('GIT_ID'),":",Sys.getenv('GIT_TOKEN'),"@gitlab.met.no/pierreml/sitingclass.git"),force=T)

.rs.restartR()
library(sitingclass)
```

### Authentication
Met's Frost API is used to retrieve the necessary station metadata including coordinates and station name. First [Register as a user](https://frost-beta.met.no/docs/starthere) to obtain `YOUR_FROST_ID` and `YOUR_FROST_KEY`. To make them accessible to your R environment, save them in a `.Renviron` file (more info [here](https://rstats.wtf/r-startup.html)), such as:
```{bash, eval=FALSE}
cd $HOME
cat > .Renviron
GIT_ID=YOUR_GIT_ID
GIT_TOKEN=YOUR_GIT_TOKEN
FROST_ID=YOUR_FROST_ID
FROST_KEY=YOUR_FROST_KEY
```
The function `get_metadata_frost()` will fetch `FROST_ID` and `FROST_KEY` to set `authenticate()` for the URL request.

### Dependencies
`sitingclass` depends on the following packages:

* Data and metadata retrieval:
  + [httr](https://httr.r-lib.org/)/[httr2](https://httr2.r-lib.org/): to get station metadata using **URL get request** (i.e. Frost)
  + [maptiles](https://github.com/riatelab/maptiles): to get tiles from available map provides
  + [ows4R](https://github.com/eblondel/ows4R): to download DEMs using **WCS getCoverage** (Web Coverage Service)
* Geospatial processing:
  + [sf](https://github.com/rspatial/sf): vector processing and conversion
  + [terra](https://github.com/rspatial/terra): SpatRaster and SpatVector processing and conversion
* Computing horizon and sun position:
  + [rgrass](https://rsbivand.github.io/rgrass/): to compute horizon from GRASS
  + [suntools](https://github.com/adokter/suntools/): to compute sun position in the sky
* Plotting:
  + [ggplot2](https://ggplot2.tidyverse.org/): plotting functions
  + [rayshader](https://www.rayshader.com/): to plot DEMs with shaded relief and in 3D
  + [tidyterra](https://dieghernan.github.io/tidyterra/): to plot raster images and contour
* Utils:
  + [magrittr](https://magrittr.tidyverse.org/): pipping (i.e. `%>%`)
  
### Data and metadata 
`sitingclass` aims to be portable retrieving most of the needed data sets from public web services, including:

- [Frost](https://frost-beta.met.no/docs/codeexamples): station metadata
- [Kartverket](https://www.kartverket.no/geodataarbeid/nasjonal-detaljert-hoydemodell)/[Geonorge](https://kartkatalog.geonorge.no/metadata?text=25833+WCS+h%C3%B8ydemodell): digital terrain model and digital surface model in UTM 33 (epsg:25833)
- [Nibio](https://nibio.no/tjenester/wms-tjenester/wms-tjenester-ar5): WMS (Web Map Services) Land cover maps

## Usage

### Get the station's metadata

```{r}
library(sitingclass)

# Define station id
stationid <- 18700

# Get station metadata
stn <- get_latlon_frost(stationid)
stn
```
### Construct the coordinate box
```{r}
# Get coordinates
centre <- sf::st_coordinates(stn)
centre

# Construct box to extract WMS tile
dx <- 100
box <- round( c(c(centre[1],centre[2])-dx,
                c(centre[1],centre[2])+dx) )
class(box) <- "bbox"
box <- sf::st_as_sfc(box)
sf::st_crs(box) <- 25833 #32633 #to match tile projection

```

```{r,fig.dim = c(10, 8), out.width="80%", fig.cap="A satellite image with the station location as a red dot."}
# Plot ESRI imagery
plot_tile_station(stn, box, tile_name="esri")
```

### Download the DEMs
```{r,fig.dim = c(10, 8), out.width="80%", fig.cap="Hillshade of the digital surface model.", results = 'hide'}
f.ow  <- FALSE
dem   <- download_dem_kartverket(stationid,centre,name="dtm",dx,resx=1,f.overwrite=f.ow)
dsm   <- download_dem_kartverket(stationid,centre,name="dom",dx,resx=1,f.overwrite=f.ow)
demkm <- download_dem_kartverket(stationid,centre,name="dtm",20e3,resx=20,f.overwrite=f.ow)

# Plot 100-m Digital Surface Model and 20-km Digital Terrain Model
library(rayshader)
elmat   = raster_to_matrix(dsm)
elmatkm = raster_to_matrix(demkm)

#We use another one of rayshader's built-in textures:
elmat %>%
  sphere_shade(texture = "desert") %>%
  plot_map()

elmatkm %>%
  sphere_shade(texture = "desert") %>%
  plot_map()
```
### Compute and plot the horizon and sun diagram
```{r,fig.dim = c(8, 6), out.width="80%", fig.cap=paste("Horizon and sun diagram for the station ",stationid)}
plot_station_horizon_sun(stn, dem, dsm, demkm)
```

## `plot_station_siting_context()`
The function `plot_station_siting_context()` will compute all the code chunks from [Usage] such as:

```{r}
# Set timezone to avoid time shift between winter and summer time
Sys.setenv(TZ="UTC") # "Europe/Oslo"

# The main function plotting sun diagram and context for a weather station 
plot_station_siting_context(stationid=18700, paramid=211, f.verbose=F, f.pdf=F)

```


